---
title: "Week 3 - Lab Exercise"
author: "Garen Bainbridge"
date: "2025-09-03"
output: html_document
---
```{r load tidyverse packages}
library(tidyverse)
```

```{r}
socialMedia <- read_csv("socialMediaAddiction.csv")
glimpse(socialMedia)
```
```{r histogram_age}

socialMedia %>%
  ggplot(mapping = aes(x = Age)) +
  geom_histogram()
```
The distribution of the age variable is multimodal with 4 distinct peaks and not symmetric. I hypothesize that most of the individuals in this dataset are college students, which explains the clustering of peaks around the ages 19-22 in the histogram.
```{r histogram_Avg_Daily_Usage_Hours}
socialMedia %>%
  ggplot(mapping = aes(x = Avg_Daily_Usage_Hours)) +
  geom_histogram()
```
The distribution of the Avg_Daily_Usage_Hours variable is unimodal with 1 distinct peak and not symmetric. I hypothesize that the peak and cluster between 4-6 hours is due to the participants being college students that use social media very often.

To visualize the relationship between students who have higher social media usage and getting fewer hours of sleep, a scatterplot is the best type of graph to use. The variable on the x-axis is Avg_Daily_Usage_Hours, and the variable on the y-axis Sleep_Hours_Per_Night.

```{r scatterplot}
socialMedia %>%
  ggplot(mapping = aes(x = Avg_Daily_Usage_Hours, y = Sleep_Hours_Per_Night))+
  geom_point()
```
Social media usage and sleep seem to have a clear negative correlation.

A boxplot would be best to visualize the relationship between longer social media usage with specific apps. The x-axis would represent the different social media apps being analyzed, and the y-axis would represent the average daily usage hours.

```{r boxplot}
socialMedia %>%
  ggplot(mapping = aes(x = Most_Used_Platform, y = Avg_Daily_Usage_Hours)) + 
    geom_boxplot()
```

These boxplots suggest a relationship between social media platform and daily usage hours with Facebook, Instagram, Tiktok, and WhatsApp having the highest average daily usage hours.

Does the relationship between social media usage and mental health vary by the age of the student?

A scatterplot would best visualize the relationship. The x-axis would represent average daily social media usage, and the y-axis would represent mental health score. The feature used for age, the third variable, is facet_wrap() to create separate scatterplots for each age.

```{r scatterplot}
socialMedia %>%
  ggplot(mapping = aes(x = Avg_Daily_Usage_Hours, y = Mental_Health_Score))+
  geom_point()+
  geom_smooth(method ="lm")+
  facet_wrap(~Age)
```

There is a negative correlation between average daily social media usage and mental health score.

NEXT STEPS SECTION

```{r raster plot}
socialMedia %>%
ggplot(aes(x = Avg_Daily_Usage_Hours,
                        y = Mental_Health_Score,
                        fill = Age)) +
  geom_raster(hjust = 0.5, vjust = 0.5, interpolate = FALSE) +
  labs(title = "Raster Plot of Social Media Usage vs Mental Health, Colored by Age",
       x = "Average Daily Usage (Hours)",
       y = "Mental Health Score",
       fill = "Age")
```
Does the relationship between average daily social media usage and addicted score vary depending on how many conflicts they have over social media?

```{r raster plot}
socialMedia %>%
ggplot(aes(x = Avg_Daily_Usage_Hours,
                        y = Addicted_Score,
                        fill = Conflicts_Over_Social_Media)) +
  geom_raster(hjust = 0.5, vjust = 0.5, interpolate = FALSE) +
  labs(title = "Raster Plot: Social Media Usage vs Feelings of Addiction, Colored by Number of Social Media Conflicts",
       x = "Average Daily Usage (Hours)",
       y = "Addicted Score",
       fill = "Social Media Conflicts")
```
The raster plot suggests that students who have a higher social media usage also feel more addicted and are more prone to social media conflicts. I chose a raster plot to visualize this relationship as it allows me to display three different variables in a 2D space. The x and y axes represent the relationship between social media usage and feelings of addiction, while the color adds a third variable, indicating social media conflicts overlap with the x-y relationship.